---
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm-latex-ms_REV.tex
title: "[ECO723] Assignment #1"
author:
- name: Daeyoung Lim
  affiliation: Dept. of Statistics, Korea University
header-includes:
- \usepackage{cancel}
- \usepackage{amsmath,amssymb}
- \usepackage{graphicx}
- \usepackage{bm}
#abstract: "This document provides an introduction to R Markdown, argues for its benefits, and presents a #sample manuscript template intended for an academic audience. I include basic syntax to R Markdown and a #minimal working example of how the analysis itself can be conducted within R with the `knitr` package."
#keywords: "pandoc, r markdown, knitr"
date: "2017-09-20"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
#biblio-style: apsr
---

# Question #3.1

## Answer
If we let
$$
\begin{aligned}
X_{1},\ldots, X_{n} &\overset{\text{iid}}{\sim} N\left(0,1\right)I(X \geq 0)\\
Y_{1},\ldots, Y_{n} &\overset{\text{iid}}{\sim} N\left(0,1\right)I(Y < 0)
\end{aligned}
$$
One way of sampling from a truncated normal distribution $N(\mu, \sigma^{2})I(X \in [\alpha, \beta])$ is by *the probability integral transform* or the *inverse CDF method*. That is,
$$
X = \Phi^{-1}(\Phi(\alpha)+U(\Phi(\beta)-\Phi(\alpha)))\sigma + \mu
$$
where $U\sim \operatorname{Unif}(0,1)$.
```{r plotTN, fig.height = 3, fig.align = 'center'}
library(latex2exp)
n = 30000L # sample size
t1 = pnorm(0, mean = 0, sd = 1)
t1[t1 > (1 - 1.0e-07)] = 1 - 1.0e-07
u1 = runif(n, min = 0, max = 1)
x = qnorm(t1 + u1 * (1 - t1), mean = 0, sd = 1)

t0 = pnorm(0, mean = 0, sd = 1)
t0[t0 < 1.0e-10] = 1.0e-10
u2 = runif(n, min = 0, max = 1)
y = qnorm(t0 * u2, mean = 0, sd = 1)
par(mar=c(3.5,3.5,1.5,1), mgp=c(2.4,0.8,0), las=1, mfrow=c(1,2))
hist(x, ylab = "", probability = TRUE, breaks = 20,
     main = TeX("Histogram of TN$_{\\[0, \\infty)}$"))
curve(2 * dnorm(x), from = 0, to = max(x), add = TRUE, col = "red", lwd = 2)
hist(y, ylab = "", probability = TRUE, breaks = 20,
     main = TeX("Histogram of TN$_{(-\\infty, 0\\]}$"))
curve(2 * dnorm(x), from = 0, to = min(y), add = TRUE, col = "red", lwd = 2)
```

# Question #3.2

## Answer
If $f(x) = x / 2$,
$$
\begin{aligned}
F(x) &= \int_{0}^{x}\dfrac{u}{2}\,\mathrm{d}u = \dfrac{x^{2}}{4},\quad 0\leq x \leq 2\\
F^{-1}(u) &= \sqrt{4u},\quad 0\leq u \leq 1
\end{aligned}
$$
Therefore, by the *probability integral transform*,
$$
X = F^{-1}(U) = \sqrt{4U},\quad U\sim \operatorname{Unif}(0,1)
$$
Analytically, the expected value of $X$ is
$$
\operatorname{E}(X) = \int_{0}^{2}\dfrac{x^{2}}{2}\,\mathrm{d}x = \left.\dfrac{x^{3}}{6}\right|_{0}^{2} = \dfrac{4}{3}
$$
Sampling $X_{1},\ldots,X_{n} \overset{\text{iid}}{\sim} F$, we can use the law of large numbers to approximate $\operatorname{E}(X)$, i.e.,
$$
\dfrac{1}{n}\sum_{i=1}^{n}X_{i}\xrightarrow{n\to\infty}\operatorname{E}(X)
$$
```{r}
n = 30000L
u = runif(n, min = 0, max = 1)
x = sqrt(4 * u)
Ex = mean(x)
cat("E(X) = ", 4/3, "\n",
    "Xbar = ", Ex, sep = "")
```

# Question #3.3

## Answer
We know that $X \sim \operatorname{Laplace}(0,1)$. The inverse CDF (or the quantile function) of $\operatorname{Laplace}(0,1)$ is
$$
F^{-1}(u) = -\operatorname{sgn}\left(u-\dfrac{1}{2}\right)\log\left(1-2\left|u-\dfrac{1}{2}\right|\right)
$$
Therefore, we can plug in $U\sim \operatorname{Unif}(0,1)$ to generate random variables from $\operatorname{Laplace}(0,1)$.
```{r plotExp, fig.width=3.5, fig.height=3, fig.align='center'}
n = 20000L
u = runif(n, min = 0, max = 1)
DE = -sign(u - 0.5) * log(1 - 2 * abs(u - 0.5))
par(mfrow = c(1,1), mar=c(3.5,3.5,1,1), mgp=c(2.4,0.8,0), las=1)
hist(DE, ylab = "", probability = TRUE, breaks = 20,
     main = "Histogram of Laplace(0,1)")
curve(0.5 * exp(-abs(x)), from = min(DE), to = max(DE), add = TRUE, col = "red", lwd = 2)
```

# Question #3.4

## Answer
If $X \sim f$ is sought but a random variable with density $g$ is much easier to generate, we can find $\beta$ to satisfy a bound on the densities
$$
\beta f(x)\leq g(x)\quad \text{for }0<\beta<1
$$
Since we are dealing with a mathematical optimization problem, we can do without all the constants multiplied, which reduces the objective function to
$$
\hat{x} =\arg\min_{x\in \mathbb{R}} \dfrac{\exp\left(-(x-1/2)^{2}/2\right)}{x^{2}(1-x)^{2}}
$$
Moreover, the monotonically increasing function does not affect the minimizer of the function. Thus, applying the logarithm,
$$
\hat{x} = \arg\min_{x\in \mathbb{R}} -\dfrac{1}{2}\left(x-\dfrac{1}{2}\right)^{2} - 2 \log x - 2 \log(1-x)
$$
Then by the first-order condition, we find the root of the following equation
$$
-\left(x-\dfrac{1}{2}\right)+\dfrac{2}{1-x}-\dfrac{2}{x} = 0
$$
which gives $x = 1/2$ as the minimizer. Thus, $\hat{x}=1/2$ and $c = 2$.
Then, the acceptance-rejection (AR) algorithm goes by

* Generate $X$ from $g$.
* Generate $U$ from $\operatorname{Unif}(0,1)$.
* If $Ug(X) \leq \beta f(X)$ then accept $X$; else reject and repeat from the beginning.
```{r plotTN01, fig.width=3.5, fig.height=3, fig.align='center'}
library(truncnorm)
n = 50000L
u = runif(n, min = 0, max = 1)
tx = rtruncnorm(n, a = 0, b = 1, mean = 0.5, sd = 1)
u = runif(n, min = 0, max = 1)
g = dtruncnorm(x = tx, a = 0, b = 1, mean = 0.5, sd = 1)
f = dbeta(tx, 3, 3)
M = dtruncnorm(0.5, a = 0, b = 1, mean = 0.5, sd = 1) / dbeta(0.5, 3, 3)
tx = tx[u * g < M * f]
par(mfrow = c(1,1), mar=c(3.5,3.5,1,1), mgp=c(2.4,0.8,0), las=1)
hist(tx, ylab = "", probability = TRUE, breaks = 20,
     main = TeX("Histogram of TN$_{(0,1\\]}(0.5, 1)$"))
curve(dbeta(x, 3, 3)
      , from = 0, to = 1, add = TRUE, col = "red", lwd = 2)
cat("Var(X) = ", 9 / (6^2 * 7), "\n",
    "MCvar(X) = ", var(tx), sep = "")
```

# Question #3.6

## Answer
To get the minimizer, we need to solve the following optimization problem:
$$
\hat{x} = \arg\min_{x\in \mathbb{R}} \dfrac{x^{2}}{2}-|x|
$$
By differentiating the objective function with respect ot $x$,
$$
x - \dfrac{x}{|x|} = 0
$$
We get $x = -1$ or $x = 1$ as the minimizers. Thus, $\hat{x} = -1$ or $1$ and $\beta = \sqrt{\pi / (2e)}$ which means $c = \sqrt{2e / \pi}$. Thus,
```{r plotSNcompare, fig.width=3.5, fig.height=3, fig.align='center'}
par(mfrow = c(1,1), mar=c(3.5,3.5,1,1), mgp=c(2.4,0.8,0), las=1)
n = 50000L
u = runif(n, min = 0, max = 1)
DE = -sign(u - 0.5) * log(1 - 2 * abs(u - 0.5))

u = runif(n, min = 0, max = 1)
M = 0.5 * exp(-1) / dnorm(1)
stdNormal = DE[u * (0.5 * exp(-abs(DE))) <= M * dnorm(DE)]
hist(stdNormal, probability = TRUE,
     main = "Histogram of Standard Normal", ylab = "", xlab = "x")
curve(dnorm(x), from = min(stdNormal), to = max(stdNormal),
      add = TRUE, col = "red", lwd = 2)
cat("E(x) = ", mean(stdNormal), "\n",
    "Var(x) = ", var(stdNormal), sep = "")
```